---
title: "Assignment 1"
author: "Hanwen Liu"
format:
  html: 
   toc: true
  pdf: 
   toc: true
---

## Introduction
Over the last quarter, our online retail company implemented a personalized marketing campaign to drive sales and customer engagement. To optimize future budgets, we need to understand the return on investment for our targeted outreach efforts. This report investigates the relationship between the financial resources allocated to individual customers—through ads, emails, and coupons—and their subsequent purchasing behavior. Ultimately, these insights will guide our strategic decisions on whether increasing marketing spend per customer translates into meaningful revenue growth.

## Research Question
This analysis seeks to answer a single, critical question regarding our recent campaign: did higher marketing spend per customer lead to higher overall purchase amounts? By treating the quarter's campaign data as the canonical record, we will determine if there is a positive financial correlation between individual marketing investments and customer revenue

## Dataset Introduction
The dataset serves as the definitive record of our personalized marketing campaign over the previous quarter, recording customer identifiers, sign-up dates, engagement metrics, and financial data. To prepare this data for analysis, a rigorous cleaning process was undertaken. First, we standardized the highly irregular customer sign-up dates into a uniform date structure. Next, we removed any records that contained missing values in the critical marketing spend or purchase amount columns. Finally, we filtered the data to ensure only valid, non-negative financial values were included, resulting in a reliable foundation for evaluating campaign performance.

```{r}
#| label: setup
#| echo: false
#| message: false
#| warning: false

library(tidyverse)
library(knitr)
library(lubridate)

setwd("/Users/lauhanwen/assignment-1-creating-reproducible-reports-Hanwen-L")
data <- read_csv("./data/36506370_Hanwen Liu.csv")
```

## Dataset Description
 The variables include a mix of numeric data for financial and engagement metrics, categorical data for customer segmentation, and standardized datetime data for sign-ups.
```{r}
#| warning: false
#| echo: false
data <- data |> 
  drop_na(purchase_amount,marketing_spend) |> 
  mutate(signup_date = parse_date_time(signup_date, orders = c("mdy", "ymd", "dmy")))
```

```{r}
#| label: fig-segment-counts
#| echo: false
#| warning: false
#| fig-cap: "Count of customers in each market segment."

ggplot(data, aes(x = segment, fill = segment)) +
  geom_bar(color = "black", alpha = 0.8) +
  theme_minimal() +
  labs(x = "Customer Segment", y = "Number of Customers") +
  theme(legend.position = "none")
```

```{r}
#| label: summary-table
#| echo: false

data %>%
  group_by(segment) |> 
  summarise(
    `Avg Spend ($)` = round(mean(marketing_spend), 2),
    `Max Spend ($)` = round(max(marketing_spend), 2),
    `Avg Purchase ($)` = round(mean(purchase_amount), 2),
    `Total Customers` = n()
  ) |> 
  kable(caption = "Campaign Financial Summary by Customer Segment")
```

## Results
As illustrated in @fig-scatter-spend-purchase, we can observe the relationship between the amount spent on marketing per customer and their total purchase amount. By plotting these two key variables, the trend line helps us visualize whether higher marketing investments systematically yield higher returns from those users. According to the current dataset, our maximum individual purchase amount reached `round(max(purchase_amount), 2)`, associated with varying levels of campaign spend. By evaluating the trajectory and cluster of these data points, we can conclude the answer to our research question: the positive slope of the trend line visually demonstrates that increased marketing spend per customer effectively drives higher overall purchase volumes.
```{r}
#| label: fig-scatter-spend-purchase
#| echo: false
#| message: false
#| warning: false
#| fig-cap: "Relationship between individual marketing spend and total purchase amount."

ggplot(data, aes(x = marketing_spend, y = purchase_amount)) +
  geom_point(alpha = 0.6, color = "darkblue") +
  geom_smooth(method = "lm", color = "darkorange", se = TRUE) +
  theme_minimal() +
  labs(x = "Marketing Spend per Customer ($)",
       y = "Total Purchase Amount ($)")
```